---
title: "Project 2B: Streaming Data"
author: "Cameron Pittman"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: cosmo
    highlight: textmate
    code_folding: hide
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```
***

## Setup

### Packages Used

```{r loadpackages}
library(tidyverse)
library(readr)
library(waffle)
library(ggbeeswarm)
```

### Datasets Used

### Background of the Data

The datasets were downloaded from [Disney+ Movies and TV Shows | Kaggle](https://www.kaggle.com/datasets/shivamb/disney-movies-and-tv-shows) and [TV shows on Netflix, Prime Video, Hulu and Disney+ | Kaggle](https://www.kaggle.com/datasets/ruchi798/tv-shows-on-netflix-prime-video-hulu-and-disney). You can read about them there, including variable definitions, sources, when they were created, and other information. Load the two datasets and use `glimpse()` to explore their structures.

#### Disney+ Dataset

```{r loaddisneydata}
#sets the working directory to load data
setwd('/home/students/pittmancr/git_repos/STT2860S22project2/data')
#loads the data from disneypluscontent and names the dataset disney_data
disney_data <- read_csv("disneypluscontent.csv")
#displays the contents of streaming_data
glimpse(disney_data)
```

#### Streaming Dataset

```{r loadstreamingdata}
#sets the working directory to load data
setwd('/home/students/pittmancr/git_repos/STT2860S22project2/data')
#loads the data from streamingcontent and names the dataset streaming_data
streaming_data <- read_csv("streamingcontent.csv")
#displays the contents of streaming_data
glimpse(streaming_data)
```

<hr>

## Analyses

### Create the Analysis Datasets

#### Dataset #1

The Disney+ dataset contains a variable that gives information about the `rating` for each program offering. Convert the variable into a factor that puts the ratings in (roughly) ascending order of age restriction/audience: `TV-Y`, `TV-Y7`, `TV-Y7-FV`, `TV-G`, `G`, `TV-PG`, `PG`, `PG-13`, `TV-14`. If the `date_added` variable is not already in date format, convert it.  

```{r factorratings}
disney_data$rating <- factor(disney_data$rating, levels = c("TV-Y", "TV-Y7", "TV-Y7-FV", "TV-G", "G", "TV-PG", "PG", "PG-13", "TV-14"))
disney_data
```

#### Dataset #2

From the streaming content dataset, create a new dataset by doing the following.

* Use `pivot_longer()` to gather the `Netflix`, `Hulu`, `Prime Video`, and `Disney+` variables into columns called `Service` (containing the name of the streaming service) and `YesNo` (showing the 0/1 indicator values from the original columns).
* After pivoting, filter the dataset to remove any row where `YesNo == 0` (i.e., a show is *not* on the service).
* Split the `IMDb` variable to separate the show rating from the max rating of `10`. Convert the show rating to a numerical variable.
* Split the `Rotten Tomatoes` variable to separate the show rating from the max of `100`. Convert the show rating to a numerical variable.

```{r pivotlong}
streaming_datalong <- streaming_data %>%
  pivot_longer(
    cols = `Netflix`:`Disney+`,
    names_to = "Service",
    values_to = "YesNo"
  ) %>%
  filter(YesNo == 1) %>%
  separate(
    col = IMDb,
    into = c("IMDb", "maxIMDb"),
    sep = "/",
    convert = TRUE
           ) %>%
  separate(
    col = `Rotten Tomatoes`,
    into = c("Rotten Tomatoes", "maxRotten Tomatoes"),
    sep = "/",
    convert = TRUE
           )
streaming_datalong
```

### Visualization 1: Violin Plot

**This plot uses Dataset #1.**

Create a violin plot of `release_year` (x-axis) grouped by `type` of program (y-axis) for programs on Disney+. Fill with a color of your choice. Add a boxplot inside the violin plot, as you did in one of the DataCamp exercises. Re-scale the x-axis so that tick marks appear at whole-decade intervals (e.g., 1980, 1990, 2000). Add an appropriate title and axis labels. Use other formatting as you choose to enhance effectiveness/appearance.

```{r violinplot}
disney_data %>%
  ggplot(aes(x = release_year, y = type)) +
  geom_violin(fill = '#113CCF', alpha = 0.6, bw = 1) +
  geom_boxplot(fill = "#BFF5FD", alpha = 0.4, width = 0.4) +
  scale_x_continuous(breaks = c(seq(from = 1910, to = 2030, by = 10))) +
  labs(title = 'Type of Program by Release Year', y = 'Type of Program', x = 'Release Year')
```

### Visualization 2: Waffle Plot

**This plot uses Dataset #1.**

Create a waffle plot to show the distribution of program `type` on Disney+. 

* Give the plot the title "Streaming Content on Disney+". 
* Change the colors of the squares to something other than the defaults.
* Use an x-axis label to indicate roughly how many programs each square represents.

Hint: Use `round(100 * prop.table(table(DATASETNAME$VARIABLENAME)))` to create the input data for the waffle plot. Swap out the capital letter placeholders for the correct dataset name and variable name.

```{r waffleplot}
type_percent <- round(100 * prop.table(table(disney_data$type)))

waffle(type_percent, colors = c("#113CCF", "#BFF5FD")) +
  labs(title = "Streaming Content on Disney+", x = "Each square represents roughly 137 programs")
```

### Visualization 3: Ratings Bar Plot

**This plot uses Dataset #1.**

Create a horizontal barplot to show the frequencies of the various `rating` categories for programs on Disney+ (i.e., use `y =` rather than `x =` for the variable you are plotting). **You will need to remove a couple `NA` values first.** Facet by `type` so that you have side-by-side plots for the two different program types. Add an appropriate title and axis labels, as well as other formatting to enhance effectiveness/appearance. 

```{r ratingsbarplot}
disney_data_omitna <- disney_data[-c(194, 198),]
  ggplot(disney_data_omitna, aes(y = rating,)) +
  geom_bar(fill = "#113CCF") +
  facet_wrap(~type) +
  labs(title = "Program Rating Counts by Program Type", x = "Number of Programs", y = "Rating")
```

### Visualization 4: Services Bar Plot

**This plot uses Dataset #2.**

Create a barplot to display how many shows are offered on each of the four streaming services. Choose appropriate colors, labels, themes, and/or and other types of formatting that you feel will enhance the meaning or visual appearance of the plot. 

```{r showsbarplot}
streaming_datalong %>%
  ggplot(aes(x = Service, fill = Service)) +
  geom_bar() +
  scale_fill_manual(values = c("#113CCF",
                               "#66aa33",
                               "#E50914",
                               "#00A8E1")) +
  labs(title = "Programs by Streaming Service", x = "Streaming Service", y = "Number of Programs")
```

### Visualization 5: Choose a Plot!

Create one other plot of your choice using either dataset to explore a question of interest. You are welcome to perform additional manipulations on the data, if needed. Add an appropriate title and axis labels, as well as any other necessary formatting.

```{r IMDbeeswarm}
streaming_datalong %>%
  filter(Service == c("Netflix", "Hulu", "Prime Video")) %>%
  ggplot(aes(x = Service, y = IMDb, color = Service)) +
  geom_beeswarm(cex = 1, alpha = 0.8) +
  labs(title = "Top Three Streaming Service IMDb Ratings", x = "Streaming Service", y = "IMDb Rating (out of 10)")
```

<hr>

## Questions

**Question 1:** Based on your plots, make five informational statements or comparisons regarding the Disney+ streaming service.

**ANSWER**

1) Disney+ has fewer programs available than any of the other streaming services in these datasets.

2) Most movies available on Disney+ are rated between TV-G and PG.

3) The Disney+ platform is approximately 72% movies.

4) Most of the TV Shows available on Disney+ were released after 2010.

5) The ratings of the TV shows available on Disney+ are more variable than those of the movies available on the platform.

**Question 2:** What other data would you like to have, or which existing variables would you like to see transformed, if you were going to do further explorations or visualizations? Give at least two examples.

**ANSWER**

I would like to see plots for the other three streaming services of type of program by release year, similar to the plot we created for Disney+. I would also like to see transformed data based on the updated ratings guidelines compared to those of older films and tv shows, as the rating system has changed over the years. That might apply more to the Disney+ data set, since they have all of the old Disney cartoons available.

**Question 3:** Explain the rationale behind the choices you made with regard to plot type, formatting, and so on, when you created Visualization 5. What motivated your decisions? 

**ANSWER**

I was curious about how different platforms' content compared to each other in terms of IMDb ratings. I thought that a beeswarm plot would be a sufficient vizualisation for that question, and I wanted to get more practice creating beeswarm plots. I attempted to make the colors of the points align with the brand colors of the services like I was able to for the barplot, but I was unable to get that to work. It was frustrating because while Hulu and Netflix's brand colors are green and red respectively, the plot flipped those colors. 

***
```{r}
sessionInfo()
```
